---
title: "LMH ADCtraining Webpage"
author: "Leslie M. Hartten"
date: "10/10/2019"
output: 
  html_document: 
    toc: yes
---
# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries (Packages)

```{r package_load, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT) #install.packages("DT")
library(leaflet)
```

# Read in and clean up data

```{r data_load, results='hide'}

# Only read data if don't have it
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
#    read.csv("data/escapement.csv", stringsAsFactors = FALSE),
    read.csv("data/ADFG_firstAttempt_reformatted.csv", stringsAsFactors = FALSE),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
        return(esc)
    }
)

head(esc)
```

# Calculate Annual Escapement for Salmon Species

```{r}
species_list <-  c("Chinook", "Sockeye", "Chum", "Pink", "Coho")

# Extract Year info
# Group by Species, Region, Year
#   group_by order affects table orders but not values
# Calculate (Summarize) annual from daily data
# Focus on a few species
annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-", remove = FALSE) %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Year, SASAP.Region, Species) %>% 
  summarise(total_escapement = sum(DailyCount)) %>% 
  filter(Species %in% species_list)

head(annual_esc)
```

# Visualize Escapement

```{r, echo=FALSE}
# stripes caused by geom_col don't show in jpg
ggplot(data = annual_esc, mapping = aes(x = Species, y = total_escapement)) +
  geom_col(fill = "blue")

ggplot(data = annual_esc, mapping = aes(x = Species, y = total_escapement, fill = SASAP.Region)) +
  geom_col()

ggplot(data = annual_esc, mapping = aes(x = Species, y = total_escapement, fill = Year)) +
  geom_col() +
  ylab("Escapement (number of fish)") +
  theme_bw()

```

```{r}
# set a theme.  theme_* are "global" for a plot, so order of use is important
my_theme <- theme_minimal() +
  theme(legend.position = "bottom")

# plot, using the stuff in my_theme
ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"),
       mapping = aes(x = Year, y = total_escapement, color = Species)) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  ylab("Escapement (number of fish)") +
  ggtitle("Kodak Salmon Escapement") +
  my_theme
```

```{r}
ggplot(annual_esc, aes(x = Year, y = total_escapement, color = Species)) +
  geom_point() +
  geom_line() +
  facet_wrap(~SASAP.Region, scales = "free_y") +
  my_theme

ggsave("figures/escapementRegions.png")
```

```{r, fig.width = 7, fig.height = 15}
fig1 <- ggplot(annual_esc, aes(x = Year, y = total_escapement, color = Species)) +
  geom_point() +
  geom_line() +
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 1)

fig1

# Device to use. Can either be a device function (e.g. png()), or one of "eps", "ps", "tex" (pictex), "pdf", "jpeg", "tiff", "png", "bmp", "svg" or "wmf" (windows only).
ggsave("figures/escapementRegions2.png", fig1, width = 12, height = 7, units = "in")
```

# Map Sampling Locations

Create a table with unique sampling locations, latitudes, and longitudes.
```{r}
locations <-  esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()

head(locations)
```

Create table of locations and positions.  Resulting html table is very, very slick!
(Note, separated chunk that creates output from chunk that did analysis.)
```{r}
datatable(locations)
```

```{r}
# could use mutate call to create/define popup text

leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lat = ~Latitude, lng = ~Longitude, popup = ~Location)

# Projection is from the projection of tiles (set for open street maps & google street maps).  In options, can change projection, but then no longer use those tiles!  Need GIS layers or some other built basemap.
# SF has projection options
# Maybe ADC will build ...
# Tiles are not required.  You could provide a shapefile for coasts, etc.
```

```{r}
leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?",
              layers = 'GEBCO_LATEST',
              attribution = "Imagery reproduced from the GEBCO_2014 Grid, version 20150318, www.gebco.net") %>%
  addCircleMarkers(lng = ~Longitude,
                   lat = ~Latitude,
                   popup = ~ Location,
                   radius = 5,
                   # set fill properties
                   fillColor = "salmon",
                   fillOpacity = 1,
                   # set stroke properties
                   stroke = T,
                   weight = 0.5,
                   color = "white",
                   opacity = 1)
```




