---
title: "Our Best ADC Results"
author: ["Leslie M. Hartten", "Jeanette Clark", "Matthew B. Jones", "Edward Correlation"]
date: "10/11/2019"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library_load, echo=FALSE}
library(ggplot2)
library(broom) # install.packages("broom")
library(knitr)
library(captioner)
```

```{r data_load, echo=FALSE, message=FALSE, results='hide'}
# stuff won't show up in knit document
data(diamonds)
head(diamonds)
```

# Abstract

This is my abstract.  It should have a lot of words, but probably no more than 250.

# Introduction

This is the start of my introduction.  It is meet and good to cite other groups like R [@RCoreTeam] using citations in RMarkdown.

Writing reports and academic papers is a ton of work but a large amount of that work can be spent doing monotonous tasks such as:

- Updating figures and tables as we refine our analysis
- Editing our analysis and, in turn, editing our paper's text
- Managing bibliography sections and in-text citations/references

These monotonous tasks are also highly error-prone.
With RMarkdown, we can close the loop, so to speak, between our analysis and our manuscript because the manuscript can become the analysis.

As an alternative to Microsoft Word, RMarkdown provides some advantages:

- Free to use
- Uses text so we can:
  - Use version control for
    - Tracking changes
    - Collaborating
  - Edit it with our favorite and most powerful text editors
  - Use the command line to for automation

The rest of this document will show how we get some of the features we need such as:

- Attractive typesetting for mathematics
- Figures, tables, and captions
- In-text citations
- Bibliographies

# Data and Methods

Our analysis will be pretty simple.
We'll use the `diamonds` dataset from the `ggplot2` [@ggplot] package and run a simple linear model.
At the top of this document, we started with a code chunk with `echo=FALSE` set as a chunk option so that we can load the `ggplot2` package and `diamonds` dataset without outputting anything to the screen.

The folowing construct shows the equation inline when you render it:
$$\sum_{x=1}^{N}x+1$$

For our analysis, we'll create a really great plot which really shows the relationship between price and carat and shows how we include plots in our document.
Then we'll run a linear model of the form $y = mx + b$ on the relationship between price and carat and shows how we include tables in our document.
We can also put some more advanced math in our paper and it will be beautifully typeset:

\[\sum_{i=1}^{N}{log(i) + \frac{\omega}{x}}\]

(This format shows when you hover.)

# Results

The plot we made was really great (Figure 1).

```{r pricevscarat, echo=FALSE, fig.width=8, fig.height=4, fig.align='center', cache=TRUE}
ggplot(diamonds, aes(carat, price, color = clarity)) + 
  geom_point() + # scatter plot
  labs(x = "Carat", y = "Price")
```

(Note that you could build functions for analysis, source them here, and run them.  Or you could do those things & save the results in this directory, then load them and plot them.)

Here's a little analysis.  We find that the model was even better than the data, possibly because we used a simple linear model instead of the OLS Bisector [@BabuFeigelson] method:

```{r themodel, echo=FALSE}
mod <- lm(price ~ carat, diamonds)
#kable(tidy(mod), digits = 2, caption = cap_themodel)
kable(tidy(mod), digits = 2)
```

If you have some text, you can see that our n is `r 13*4`.

We were delighted to find that the slope parameter was `r round(mod$coefficients["carat"][[1]], 2)`.


# Discussion

Incredibly insightful views ...

# References

devtools::install_github("benmarwick/rrtools")


